```{r setup, message=FALSE, include=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  "readr",
  "dplyr",
  "tidyverse",
  "rvest",
  "janitor",
  "recipes",
  "tidyr",
  "ggplot2"
)


set.seed(504)
# source("season_data.R") 
df <- read_csv("../data/player_stats.csv") %>% clean_names() %>% select(-x1)
sdf <- df 
```

## Overview

So the idea here is that we are going group NBA players into like group using a clustering algorithm The data is scraped from [Basketball Reference](https://www.basketball-reference.com/leagues/NBA_2023_per_game.html)

Lets take a look at our data.

```{r head, echo=FALSE}
df %>% head()
```

First we need to decide what columns to keep

```{r dictionary, include=FALSE}
keep <- c("player", "pos", "age", "g", "gs", "mp", "fg", "fga",
          "x3p", "x3pa", "x2p", "x2pa", "ft", "fta", "orb", "drb",
          "ast", "stl", "blk", "tov", "pts")

df <- df %>% select(all_of(keep))
```

```{r peak, echo=FALSE}
df %>% head()
```

Next we need to encode some of these categorical values.

```{r encode, echo=FALSE, warning=TRUE}
df <- df %>%  
  recipe(~ . ) %>% 
  step_dummy(pos, one_hot = TRUE) %>% 
  prep() %>% 
  bake(df)
  
```
Now let's take a look at what the distribution of the stats look like overall

```{r all_stats}
df %>% 
  pivot_longer(where(is.numeric)) %>% 
  ggplot(aes(value)) +
  geom_boxplot() +
  theme_classic() +
  facet_wrap(~name, scales = "free", nrow = 4)

```

There are some outliers here for sure but lets roll with the data as is. 

Before we model with 5 clusters (the same amount of positions on an NBA team) Lets look at the position counts as it stands today.
```{r}
sdf %>% 
  group_by(pos) %>%
  summarise(n = n()) %>% 
  ggplot(aes(n, pos, fill = pos)) +
  geom_bar(stat = "identity") +
  theme_classic()
```

```{r five_model, include=FALSE}
five_model <- df %>% 
  select(-player) %>% 
  kmeans(centers = 5)
# summary(five_model)

tidy(five_model) %>% print()
```

```{r}
df$clus_5 <- five_model$cluster
df$clus_5 <- df$clus_5 %>% as.character()

df %>% 
  group_by(clus_5) %>% 
  summarise(n = n()) %>% 
  ggplot(aes(n, clus_5, fill = clus_5)) +
  geom_bar(stat = "identity") +
  theme_classic()

df %>% 
  ggplot(aes(pts, ast, color = clus_5)) +
  geom_point() + 
  theme_classic()

  
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# df %>% 
#   group_by(clus_5) %>%
#   pivot_longer(where(is.numeric)) %>% 
#   ggplot(aes(value, fill = clus_5)) +
#   geom_density() +
#   theme_classic() +
#   facet_wrap(vars(clus_5, name), ncol = 1)

for (col in colnames(df[,2:20])) {
  # print(col)
  df$Y <- df[col] %>% unlist()
  gg_tmp <- df %>% 
    ggplot(aes(Y, fill = clus_5)) +
    geom_density() +
    theme_classic() +
    xlab(col)
  print(gg_tmp)
}
df %>% select(-Y)
```


```{r}
pct <- function(x){
  x/sum(x)
}
 
df %>%
  select(-c(pos_C:pos_SG, Y)) %>% 
  group_by(clus_5) %>%
  summarise_if(is.numeric, sum) %>% 
  summarise_if(is.numeric, pct) %>% 
  round(2) 



```













